---
title: 'Fuel efficiency : Is a manual gearbox better than an automatic?'
output: html_document
date: 'Bruno Fischer Colonimos, `r format(Sys.Date(), "%d %B %Y")`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
vsfigheight <-  2.5
vsfigwidth <-  2.5
sfigheight <-  3
sfigwidth <-  3
nfigheight <-  4
nfigwidth <-  4
lfigheight <-  5
lfigwidth <-  7
knitr::opts_chunk$set(fig.height = nfigheight, fig.width = nfigwidth)
```

```{r libs, include=FALSE}
library(knitr)
library(dplyr)
require(graphics)
library(ggplot2)
```


```{r globaldefs}

```




Executive Summary
=================
Based on a model of the fuel efficiency (mpg) of the cars in the dataframe 'mtcars', it appears that, in order to get a higher mpg, a manual gearbox is better than an automatic one for the lighter cars, and that the converse is true for the heavier cars. The neutral weight is about XXXXX pounds, a manual gearbox being about ZZZ mpg better than the same car fitted with an automatic gearbox, for a car weighing 1000 pounds less than the neutral weight. 

For a car heavier than the neutral weight, the reverse is true, an automatic gearbox providing an advantage of about ZZZ mpg over a manual gearbox, for each increase of 1000 pounds. 


Quick preliminary exploration
=============================

A quick exploration shows that, overall, in the sample, the cars fitted with a manual gearbox have a higher mpg than those with ana automatic one (mean == x-XXX for manual, vs x-YYY for automatic, see graph x-ZZZ in appendix ). However, scatterplots of mpg against each variable show , for some independent variables, a different simple regression slope for each gearbox type (see graph x-ZZZ). This suggests that things might be more complex than a single constant effect of the transmission type am, depending on what variables are in the model considered.


Model selection process
=======================

* **First approach: backwards elimination** . We start with a model with all the possible independent variables and then remove at each step the variable whith the least significantly nonzero coefficient, starting with the
highest p-values. We then refit a new model without that variable, and repeat the process until we get a model in which all the variables coefficients are significantly nonzero.\n
Using this process, we eliminate one by one x-AAA, and the resulting model formula is x-BBB, with an adjusted R² of x- CCC. This model contains the variable 'am'.

* **First diagnostic**. When plotting the residuals against the fitted values (see graph), a pattern appears, with more positive residuals at each end of the scale, and more negative residuals in the middle. This reminds us of the very different regression slopes in plot x-EEE (mpg / wt), and prompts us to add an interaction term am*wt

* **Adding interaction:** we add am*wt, thus considering the model mpg ~ am*wt + qsec. this model is better than the previous one, with an adjusted R² = xx-xxx. An anova test confirms that the additional am:wt term is a significant improvement (p.value = xx-xxx), and that adding the interaction am*qt would not be justified (p-value = xx-xxx, and all coefficients becoming non-significants)

* **Diagnostic 2: Outliers removal**: we will use the model: mpg ~ am*wt + qsec. However checking for influential outliers, we can see (plot xx-xxx) that three cars seem to modify the result. Based on this plot and the relevant dfbetas we choose to **exclude** the 'Chrysler Imperial' and the 'Maserati Bora' from the data, but to **keep** the 'Fiat 128', which has less influence on the critical coefficients (see below) $\beta_{am}$ and $\beta_{wt:am}$ than e.g. the 'Merc 240D' or the 'Cadillac Fleetwood' ($dfbeta_{am}$ = xx-xx and $dfbeta_{am:wt}$ = xx-xx)

* **Finally**: The estimated coefficients of the model are: xx-xxx
 




Using the Model
===============

* The model equation is: &ensp;&ensp; $mpg =  \beta_0 + \beta_{am} am + \beta_{hp} hp + \beta_{wt} wt + \beta_{am:wt} am \cdot wt$ &ensp;&ensp; where $\beta_0$ is the intercept and $\beta_x$ is the coefficient of x, computed from the sample. Thus, this general equation can be written as:

* if am = 1 (manual),&ensp;&ensp;&ensp;&ensp; $mpg_1 =  (\beta_0 + \beta_{am}) + \beta_{hp} hp + (\beta_{wt} + \beta_{am:wt}) wt$

* if am = 0 (automatic),&ensp;&ensp; $mpg_0 =  \beta_0 + \beta_{hp} hp + \beta_{wt} wt$
 
Thus the difference between manual and automatic is $\Delta = mpg_1 - mpg_0 = \beta_{am} + \beta_{am:wt} wt$

So manual and automatic gearbox are equivalent when $\Delta$ = 0, that is: $wt = wt_0 = -\frac{\beta_{am}}{\beta_{am:wt}}$ and the slope of the difference $\Delta$ is $\beta_{am:wt}$, i.e. a variation of weight of 1000 pounds increases $\Delta$ by  $\beta_{am:wt}$.

Point estimation: according to `lm`, $\beta_{am}$ = xx-xxx and  $\beta_{am:wt}$ = xx-yyy. thus $wt_0 \approx$ 

Interval estimation: 

* For $\beta_{am:wt}:$ we can use the standard error given by `lm` and the t distribution. We get [ ]
* For $wt_0$, we must resort to a Bootstrap procedure, and we get : [ ] (confidence level 5%)


Conclusion
==========
For a weight of about $wt_0 \approx$ xx-xxx, the expected mpg for manual and automatic gearboxes are the same.
If the car weight is 1000 pounds less than $wt_0$, the expected mpg of a car with a manual gearbox is better than that of the same car with a manual gearbox by $-\beta_{am:wt} \approx $ xx-xxx. The reverse is true if the car is 1000 pounds heavier than $wt_0$

These figures are rather rough point estimates: a 95% confidence interval for $wt_0$ is [] and for $-\beta_{am:wt}$, it is []
The relatively small sample size (...) does not allow any more accurate estimations.





Appendix
=========


figure 
--------------

(@distrib) blablabla



kjfd kj 



(@scatter) caption


figure (@)
--------------


(@) thir example
