---
title: 'Manual or automatic gearbox: Which is the best?'
output: 
  pdf_document: 
    number_sections: yes
date: "Bruno Fischer Colonimos, `r format(Sys.Date(), '%d %B %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
vsfigheight <-  2.5
vsfigwidth <-  2.5
sfigheight <-  3
sfigwidth <-  3
nfigheight <-  4
nfigwidth <-  4
lfigheight <-  5
lfigwidth <-  7
knitr::opts_chunk$set(fig.height = nfigheight, fig.width = nfigwidth)
```

```{r libs, include=FALSE}
library(knitr)
library(dplyr)
require(graphics)
library(ggplot2)
```


```{r allcode}

```

Report
======


Executive Summary 
-----------------
**Problem**: What type of gearbox is the best for improving a car's fuel efficiency (mpg)?\
**Answer**: Based on a linear model of the fuel efficiency (`mpg`) of the cars in the dataframe 'mtcars', it appears that, in order to get a higher mpg, a manual gearbox is better than an automatic one for the lighter cars, and that the converse is true for the heavier cars. The neutral weight is about XXXXX pounds, a car with a manual gearbox being about ZZZ miles per gallon better than the same car fitted with an automatic gearbox, for a car weighing 1000 pounds less than the neutral weight. 

For a car heavier than the neutral weight, the reverse is true, an automatic gearbox providing an improvement of about ZZZ miles per gallon over a manual gearbox, for each increase of 1000 pounds. 

Source document
---------------
The source (.Rmd) document for the present report can be found at:


Short preliminary exploration
-----------------------------

A quick exploration shows that, overall, in the sample, the cars fitted with a manual gearbox have a higher mpg than those with an automatic one (mean == x-XXX for manual, vs x-YYY for automatic, see graph _(@distrib)_ in appendix ). However, scatterplots of mpg against each other variable  in the set (plot _(@scatter)_ in appendix) show, for some independent variables, a different simple regression slope for each gearbox type. This suggests that, depending on what other variables are in the model considered, the influence of the gearbox type `am` on `mpg` might be more complex than a single constant effect.


Model selection process {#sec:modelselection}
-----------------------

* **First approach: backwards elimination** . We start with a model with all the possible independent variables and then remove at each step the variable whith the least significantly nonzero coefficient, starting with the
highest p-values. We then refit a new model without that variable, and repeat the process until we get a model in which all the variables coefficients are significantly nonzero.\
Using this process, we eliminate one by one x-AAA, and the resulting model formula is x-BBB, with an adjusted R² of x- CCC.  The variable of interest `am` is part of this model.

* **First diagnostic**. When plotting the residuals against the fitted values (see graph _(@diagmodel1)_), a pattern appears, with more positive residuals at each end of the scale, and more negative residuals in the middle. The same effect is observed when plotting residuals against `wt`. This reminds us of the very different regression slopes in plot _(@scatter)_ (particularly for mpg ~ wt), and prompts us to add an interaction term `am*wt`. We might also want to try to add interaction `am*qsec`.

* **Adding interaction:** we add `am*wt`, thus considering the model `mpg ~ am*wt + qsec`. This model is better than the previous one, with an adjusted R² = xx-xxx. An anova test confirms that:
    + the additional `am*wt` term is a significant improvement (p.value = xx-xxx), 
    + adding the interaction `am*qsec` would **not** be justified (p-value = xx-xxx, and all coefficients becoming non-significants). 
We thus choose the model: `mpg ~ am*wt + qsec` . The residuals plots (plot _(@diagmodel1B)_ ) now become acceptable.

* **Diagnostic 2: Outliers removal**: The model formula is `mpg ~ am*wt + qsec`. However, checking for influential outliers, we can see (plot _(@diagmodel2)_) that three cars seem to influence substantially the results. Based on this plot and on the relevant dfbetas we choose to **exclude** the 'Chrysler Imperial' and the 'Maserati Bora' from the data, but to **keep** the 'Fiat 128', which, with defbetas $dfbeta_{am}$ = xx-xx and $dfbeta_{am:wt}$ = xx-xx, has less influence on the critical coefficients $\beta_{am}$ and $\beta_{wt:am}$ (see below) than other cars ('Merc 240D', 'Cadillac Fleetwood' ...)
The same diagnostics plots after removal (plot _(@diagmodel3)_) now are acceptable.

* **Finally**: The estimated coefficients of the model are (with `lm`): 

```{r dummy}
mat <- matrix(1:25, nrow = 5)
mat <- as.data.frame(mat)
kable(mat)
```
 

Using the model to measure the influence of the gearbox type
------------------------------------------------------------

**(a) Principle**

* The model equation is: $mpg =  \beta_0 + \beta_{am} am + \beta_{qsec} qsec + \beta_{wt} wt + \beta_{am:wt} am \cdot wt$    where $\beta_0$ is the intercept and $\beta_x$ is the coefficient of x. Thus, this general equation can be written as:

    * if am = 1 (manual), $mpg = mpg_1 =  (\beta_0 + \beta_{am}) + \beta_{qsec} qsec + (\beta_{wt} + \beta_{am:wt}) wt$
    * if am = 0 (automatic), $mpg = mpg_0 =  \beta_0 + \beta_{qsec} qsec + \beta_{wt} wt$
 
Thus the difference between manual and automatic is $\Delta = mpg_1 - mpg_0 = \beta_{am} + \beta_{am:wt} wt$

So manual and automatic gearbox are equivalent when $\Delta$ = 0, that is: $wt = wt_0 = -\frac{\beta_{am}}{\beta_{am:wt}}$ and the slope of the difference $\Delta$ is $\beta_{am:wt}$, i.e. a variation of weight of 1000 pounds increases $\Delta$ by  $\beta_{am:wt}$ .


**(b) Point estimation and confidence intervals**\
As seen at the end of the [previous section](#sec:modelselection) `lm` gives us the following elements: $\beta_{am} \approx$ xx-xxx and  $\beta_{am:wt} \approx$  xx-yyy, but an estimate of the neutral $wt_0$ is not a combination of these two estimation. $wt_0$ is better estimated using a bootstrap procedure:

* For $\beta_{am:wt}:$ we can use the point estimate and the standard error given by `lm` with the t distribution. We get: 
    + point estimation: xx-yyy, 95% confidence interval [ ]
* For $wt_0$, we must resort to a Bootstrap procedure (with xxx-xx simulated samples), and we get (see graph *(@bootstrap)*): 
    + point estimation: xx-yyy (median), 95% confidence interval [ ]


Conclusion
----------
For a weight of about $wt_0 \approx$ xx-xxx, the expected mpg for manual and automatic gearboxes are the same.
If the car weight is 1000 pounds less than $wt_0$, the expected mpg of a car with a manual gearbox is better than that of the same car with a manual gearbox by $-\beta_{am:wt} \approx$ xx-xxx. The reverse is true if the car is 1000 pounds heavier than $wt_0$ .

These figures are rather rough estimates: a 95% confidence interval for $wt_0$ is [] and for $-\beta_{am:wt}$, it is []
The relatively small sample size (xx-xxx after outlier removal) does not allow a better accuracy.





Appendix: figures
=================

* (@distrib) mpg distributions


* (@scatter) Scatterplots


* (@diagmodel1) Linear model without interaction, residuals plots

* (@diagmodel1B) Linear model with interaction, residuals plots

* (@diagmodel2) Linear model with interaction, diagnostics plot for checking influential outliers.

* (@diagmodel3) Linear model  with interaction, diagnostics plot after removing two influential outliers.


* (@bootstrap) Bootstrap: Resampling distribution of $wt_0$. 



