---
title: 'Fuel efficiency : Is a manual gearbox better than an automatic?'
output: pdf_document
date: "Bruno Fischer Colonimos, 19/08/2016"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
vsfigheight <-  2.5
vsfigwidth <-  2.5
sfigheight <-  3
sfigwidth <-  3
nfigheight <-  4
nfigwidth <-  4
lfigheight <-  5
lfigwidth <-  7
knitr::opts_chunk$set(fig.height = nfigheight, fig.width = nfigwidth)
```

```{r libs, include=FALSE}
library(knitr)
library(dplyr)
require(graphics)
library(ggplot2)
```


```{r allcode}

```




Executive Summary
=================
Based on a model of the fuel efficiency (mpg) of the cars in the dataframe 'mtcars', it appears that, in order to get a higher mpg, a manual gearbox is better than an automatic one for the lighter cars, and that the converse is true for the heavier cars. The neutral weight is about XXXXX pounds, a manual gearbox being about ZZZ mpg better than the same car fitted with an automatic gearbox, for a car weighing 1000 pounds less than the neutral weight. 

For a car heavier than the neutral weight, the reverse is true, an automatic gearbox providing an advantage of about ZZZ mpg over a manual gearbox, for each increase of 1000 pounds. 

Source document
===============
The source (.Rmd) document for the present report can be found at:


Short preliminary exploration
=============================

A quick exploration shows that, overall, in the sample, the cars fitted with a manual gearbox have a higher mpg than those with an automatic one (mean == x-XXX for manual, vs x-YYY for automatic, see graph *(@distrib)* in appendix ). However, scatterplots of mpg against each other variable in the set show , for some independent variables, a different simple regression slope for each gearbox type (see graph x-ZZZ). This suggests that things might be more complex than a single constant effect of the transmission type am, depending on what variables are in the model considered.

Model selection process
=======================

* **First approach: backwards elimination**. We start with a model with all the possible independent variables and then remove at each step the variable whith the least significantly nonzero coefficient, starting with the
highest p-values. We then refit a new model without that variable, and repeat the process until we get a model in which all the variables coefficients are significantly nonzero.\
Using this process, we eliminate one by one x-AAA, and the resulting model formula is x-BBB, with an adjusted $R^2$ of x- CCC.  The variable of interest 'am'is part of this model.


* **First diagnostic**. When plotting the residuals against the fitted values (see graph (@diagmodel1)), a pattern appears, with more positive residuals at each end of the scale, and more negative residuals in the middle. The same effect is observed when plotting residuals against 'wt'. This reminds us of the very different regression slopes  in plot *(@scatter)* (particularly with mpg ~ wt), and prompts us to add an interaction term am*wt. We might also want to try to add interaction with qtime.

* **Adding interaction:** we add am*wt, thus considering the model mpg ~ am*wt + qsec. This model is better than the previous one, with an adjusted R² = xx-xxx. An anova test confirms that the additional am:wt term is a significant improvement (p.value = xx-xxx), and that adding the interaction am*qt would **not** be justified (p-value = xx-xxx, and all coefficients becoming non-significants).

* **Diagnostic 2: Outliers removal**: We thus use the model: mpg ~ am*wt + qsec. However checking for influential outliers, we can see (plot *(@diagmodel2)*) that three cars seem to substantially influence the results. Based on this plot and the relevant dfbetas we choose to **exclude** the 'Chrysler Imperial' and the 'Maserati Bora' from the data, but to **keep** the 'Fiat 128', which has less influence on the critical coefficients (see below) $\beta_{am}$ and $\beta_{wt:am}$ than e.g. the 'Merc 240D' or the 'Cadillac Fleetwood' ($dfbeta_{am}$ = xx-xx and $dfbeta_{am:wt}$ = xx-xx)


* **Finally**: The estimated coefficients of the model are: 

 
```{r}
mat <- matrix(1:9, nrow = 3)
mat <- as.data.frame(mat)
kable(mat)
```



Using the model to address the question
=======================================

### Principle {-}
* The model equation is:    $mpg =  \beta_0 + \beta_{am} am + \beta_{hp} hp + \beta_{wt} wt + \beta_{am:wt} am \cdot wt$    where $\beta_0$ is the intercept and $\beta_x$ is the coefficient of x, computed from the sample. Thus, this general equation can be written as:

* if am = 1 (manual),     $mpg_1 =  (\beta_0 + \beta_{am}) + \beta_{hp} hp + (\beta_{wt} + \beta_{am:wt}) wt$

* if am = 0 (automatic),   $mpg_0 =  \beta_0 + \beta_{hp} hp + \beta_{wt} wt$
 
Thus the difference between manual and automatic is $\Delta = mpg_1 - mpg_0 = \beta_{am} + \beta_{am:wt} wt$

So manual and automatic gearbox are equivalent when $\Delta$ = 0, that is: $wt = wt_0 = -\frac{\beta_{am}}{\beta_{am:wt}}$ and the slope of the difference $\Delta$ is $\beta_{am:wt}$, i.e. a variation of weight of 1000 pounds increases $\Delta$ by $\beta_{am:wt}$.


### Point estimation and confidence Intervals: {-} 
As seen previously `lm` gives us the following elements: $\beta_{am} \approx$ xx-xxx and  $\beta_{am:wt} \approx$  xx-yyy. but an estimate of $wt_0$ is not a combination of these 2 estimation. $wt_0$ is better estimated using a bootstrap procedure :

* For $\beta_{am:wt}:$ we can use the point estimate and the standard error given by `lm` with the t distribution. We get: 
    + point estimation: xx-yyy, confidence interval [ ] (confidence level 95%)
* For $wt_0$, we must resort to a Bootstrap procedure (with xxx-xx simulated samples), and we get (see graph *(@bootstrap)*): 
    + point estimation: xx-yyy (median), confidence interval [ ] (confidence level 95%)


Conclusion
==========
For a weight of about $wt_0 \approx$ xx-xxx, the expected mpg for manual and automatic gearboxes are the same.
If the car weight is 1000 pounds less than $wt_0$, the expected mpg of a car with a manual gearbox is better than that of the same car with a manual gearbox by $-\beta_{am:wt} \approx$ xx-xxx. The reverse is true if the car is 1000 pounds heavier than $wt_0$

These figures are rather rough point estimates: a 95% confidence interval for $wt_0$ is [] and for $-\beta_{am:wt}$, it is []
The relatively small sample size (xx-xxx after outlier removal) does not allow a better accuracy.





Appendix: figures
=================

* (@distrib) mpg distributions


* (@scatter) Scatterplots


* (@diagmodel1) Linear model diagnostic plots 1


* (@diagmodel2) Linear model diagnostic plots 2


* (@bootstrap) Bootstrap: Resampling distribution of $wt_0$




